---
title: "桑基图（Sankey）"
author: "ylchen"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = F,message = F,collapse=TRUE)
```
# 一、数据载入

* The Drug–Gene Interaction Database (DGIdb) (accessed on 8 June 2020 from http://www.dgidb.org/) is an online database of drug–gene interaction data aggregated from various sources, including several drug databases (DrugBank, PharmGKB, ChEMBL),clinical trial databases, and literature from PubMed.
* 结果为tsv文件 "dgidb_export_2021-06-04.tsv"

* 图形可分开两个数据集组成，最后合并且使用to_lodes_form 转换成长矩阵

```{r}
rm(list = ls())
#BiocManager::install("ggalluvial")
library(ggalluvial)
options(stringsAsFactors = F)
# 查看教程
#vignette(topic = "ggalluvial", package = "ggalluvial")

# 1.准备数据----
load("01.data.Rdata")
load("02_networkConstruction.RData")

dat <- exp
dat <- t(dat[order(apply(dat,1,mad), decreasing = T)[1:5408],])

genes_all <- data.frame(module=moduleColors,
                        genes=colnames(dat))
head(genes_all)

dgidb <- read.csv(file = "dgidb_export_2021-06-04.tsv",
                  comment.char = "!",
                  header = T,
                  sep = "\t")

gene <- unique(dgidb$gene)
module <- c("bule","green","brown","brown")
gene_module <- data.frame(gene,module)
gene_drug <- data.frame(gene = dgidb$gene,
                        drug = dgidb$drug)
head(gene_module)
head(gene_drug)

library(dplyr)
#合并
Tit <- inner_join(gene_drug,gene_module,by = "gene")
Tit <- Tit[,c("drug","gene","module")]
head(Tit)

# 2.改为长矩阵----
tit_long <- to_lodes_form(data.frame(Tit),
                     key = "Demographic",
                     axes = 1:3)
#目的矩阵tit_long
#三列直接对应图中
head(tit_long)

```

# 二、绘图
* https://mp.weixin.qq.com/s/Xc50nTczwcc0r8KROUU8Ig
* 交互式使用sankeynetwork包

```{r}

f7 <- ggplot(tit_long,
       aes(x = Demographic, stratum = stratum, alluvium = alluvium,
           fill=stratum, label = stratum)) +
  geom_flow() +
  geom_stratum(alpha = 3,width = 1/2) +#柱状图
  geom_text(stat = "stratum", size = 2.5) +#柱状图中标签
  theme(legend.position = "none",#去除刻度线和背景颜色
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size =15,colour = "black"),#坐标轴名
        axis.title = element_blank()) +
  scale_x_discrete(position = "top") #坐标轴位置
#ggtitle()设置图标题
f7
#ggsave(f7,filename = "fig_7.pdf")
```


